---
title: "ACICC experiment bacterial abundance"
author: "Lexie Christopoulos"
date: "10/29/2020"
output: github_document
---

#Intro 
This document shows how **individual bottle** bacterial abundance data from ACIDD experiments were processed, QD'd, and analyzed. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

#Import Data 

```{r}
excel_sheets("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx")

metadata <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx", sheet = "Metadata")
glimpse(metadata)
#unique(metadata$Experiment)
#unique(metadata$Location)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week3/ACIDD_Exp_BactAbund.xlsx", sheet = "Data")
glimpse(data)

joined <- left_join(metadata, data)
#names(joined)
#summary(joined)
glimpse(joined)

```

# Prepare Data

Convert data and time column values from characters to dates, add columns with time elapsed for each experiment, convert cells/mL to cells/L, and subset data to select only VOI and drop NA's. 


```{r}
cells <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime), cells = Cells_ml * 1000, sd_cells = Cells_ml_sd * 1000) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(interv = interval(first(Datetime), Datetime), hours = interv/3600, days = hours/24) %>% 
  ungroup() %>% 
  select(Experiment:Nutrient_Sample, hours, days, cells, sd_cells) %>% 
  drop_na(cells)

#glimpse(cells) 

```


#Plot Growth Curves 

```{r}
custom.colors <- c("Control" = "#377E88", "Ash Leachate" = "#4DAF4A", "Santa Barabara" = "#E41A1C", "San Diego" = "#FF7F00")
levels <- c("Control", "Ash Leachate", "San Diego", "Santa Barbara")

cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = cells, group = interaction(Experiment, Treatment, Bottle)))+
  geom_errorbar(aes(ymin = cells - sd_cells, ymax = cells + sd_cells, color = factor(Treatment, levels = levels)), width = 0.1)+
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1)+
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21)+
  geom_text(aes(label = dna), size = 12, color = "#E41A1C")+
  labs(x = "Days", y = expression(paste("Cells L"^-1)),fill = "Treatments")+
  guides(color = FALSE)+
  scale_color_manual(values = custom.colors)+
  scale_fill_manual(values = custom.colors)+
  facet_grid(rows = "Location", scales = "free")+
  theme_bw()
```


We can calculate: 
  - The total change in cells from the initail condition to the end of the experiment 
  - Specific growth rates as the slope of ln(abundance) v time during the exponential growth phase 
  - Doubling time as ln(2) divided by the specific growth rate 
  - The mean of each of these parameters for each treatment

First, we'll need to determine where exponential growth occurs in each of the experiments, if it does. So let's plot ln(abundance) v time. 

#Identify exponential phase of growth 

**NOTE about logarithms in R**

log(x) gives the natural log of x, not log base 10 
log10(x) gives the log base 10 
log2(x) gives log base 2

```{r}
ln_cells <- cells %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(ln_cells = log(cells), diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells))) %>% 
  ungroup()

```


```{r fig.height=4, fig.width=10}

ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>% 
  ggplot(aes(x = days, y = diff_ln_cells, group = interaction(Experiment, Treatment, Bottle)))+
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1)+
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color = "black", shape = 21)+
  geom_text(aes(label = dna), size = 12, color = "#E41A1C")+
  labs(x = "Days", y = expression(paste( "Delta ln cells L"^-1)),fill = "Treatments")+
  guides(color = FALSE)+
  scale_color_manual(values = custom.colors)+
  scale_fill_manual(values = custom.colors)+
  facet_grid(Location~Bottle, scales = "free")+
  theme_bw()

```

This plot makes it a little easier to see, with the data that we have, where exponential growth occurs for each bottle: 

-SD Bottle A ~3-5 d (T4-T6)
-SD Bottle B ~4-5 d (T5-T6)
-SD Bottle C ~2-3 d (T3-T4)
-SD Bottle D ~2-3 d (T3-T4)

-SB Bottle A ~0.5-1 d (T1-T2)
-SB Bottle B ~0.5-2 d (T1-T3)
-SB Bottle C ~1-3 d (T2-T4)
-SB Bottle D ~1-3 d (T2-T4)

#Calculating growth rates, doubling times, and Delta cell abundances 


```{r}
growth <- ln_cells %>% 
  mutate(exp_start = ifelse(Experiment == "ASH171" & Bottle == "A", 4, NA),
         exp_start = ifelse(Experiment == "ASH171" & Bottle == "B", 5, exp_start),
         exp_start = ifelse(Experiment == "ASH171" & Bottle %in% c("C", "D"), 3, exp_start),
         exp_start = ifelse(Experiment == "ASH172" & Bottle %in% c("A", "B"), 1, exp_start), 
         exp_start = ifelse(Experiment == "ASH172" & Bottle %in% c("C", "D"), 2, exp_start),
         exp_end = ifelse(Experiment == "ASH171" & Bottle %in% c( "A", "B"), 6,4),
         exp_end = ifelse(Experiment == "ASH172" & Bottle == "A", 2, exp_end),
         exp_end = ifelse(Experiment == "ASH172" & Bottle == "B", 3, exp_end)) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  mutate(ln_cells_exp_start = ifelse(Timepoint == exp_start, ln_cells, NA),
         ln_cells_exp_end = ifelse(Timepoint == exp_end, ln_cells, NA),
         cells_exp_start = ifelse(Timepoint == exp_start, cells, NA),
         cells_exp_end = ifelse(Timepoint == exp_end, cells, NA),
         days_exp_start = ifelse(Timepoint == exp_start, days, NA), 
         days_exp_end = ifelse(Timepoint == exp_end, days, NA)) %>% 
  fill(ln_cells_exp_start:days_exp_end, .direction = "updown") %>% 
  mutate(mew = (ln_cells_exp_end - ln_cells_exp_start)/(days_exp_end - days_exp_start), 
         doubling = log(2)/mew,
         delta_cells = cells_exp_end - first(cells)) %>% 
  ungroup()
         
```


#Convert bacterial abundance and change in bacterial abundance to carbon units 

Apply a carbon conversion factor (CCF) to bacterial abundance (Cells L^-1^) to bacterial carbon (micromol C L^-1^)

We'll apply the average carbon content of bacterioplankton cells from Coastal Japan (~ 30 fg C cell^-1^). This CCF was used in previous experiments conducted in the SBC. 


```{r}
bactcarbon <- growth %>% 
  mutate(bc = cells * (2.5 * 10^-9), 
         delta_bc = delta_cells * (2.5 * 10^-9))
glimpse(bactcarbon)
```


#Calculate Treatment Averages


```{r}
averages <- bactcarbon %>% 
  group_by(Experiment, Treatment, Timepoint) %>% 
  mutate(ave_bc = mean(bc), sd_bc = sd(bc)) %>% 
  ungroup() %>% 
  group_by(Experiment, Treatment) %>% 
  mutate(ave_mew = mean(mew), 
         sd_mew = sd(mew), 
         ave_doubling = mean(doubling),
         sd_doubling = sd(doubling),
         ave_delta_cells = mean(delta_cells),
         sd_delta_cells = sd(delta_cells), 
         ave_lag = mean(days_exp_start),
         sd_lag = sd(days_exp_start)) %>% 
  ungroup()

glimpse(averages)
```


#Plot Treatment Averages


```{r fig.height=4, fig.width=8}
averages %>% 
  ggplot(aes(x = days, y = ave_bc), group = interaction(Experiment, Treatment))+
  geom_errorbar(aes(ymin = ave_bc - sd_bc, ymax = ave_bc + sd_bc, color = factor(Treatment, levels = levels)), width = 0.1)+
  geom_line(aes(color = factor(Treatment, levels = levels)), size = 1)+
  geom_point(aes(fill = factor(Treatment, levels = levels)), color = "black", shape = 21, size = 3)+
  facet_grid(rows = "Location", scales= "free")+
  scale_color_manual(values = custom.colors)+
  scale_fill_manual(values= custom.colors)+
  labs(x = "Days", y = expression("Bacterial Carbon, micromol C L"^-1), fill = "")+
  guides(color = FALSE)+
  theme_bw()
```

##Barplots 


```{r}
bar.data <- averages %>% 
  select(Location, Treatment, ave_mew:sd_lag) %>% 
  distinct()
```


```{r}
mew <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_mew), group = interaction(Location, Treatment))+
  geom_col(color = "black", fill = "white")+
  geom_errorbar(aes(ymin = ave_mew - sd_mew, ymax = ave_mew + sd_mew), width = 0.1)+
  facet_grid(~factor(Location, levels = levels), scales = "free")+
  labs(x = "", y = expression("micromol, d"^-1))+
  theme_bw()
```


```{r}
doubling <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_doubling), group = interaction(Location, Treatment))+
  geom_col(color = "black", fill = "white")+
  geom_errorbar(aes(ymin = ave_doubling - sd_doubling, ymax = ave_doubling + sd_doubling), width = 0.1)+
  facet_grid(~factor(Location, levels = levels), scales = "free")+
  labs(x = "", y = expression("Doubling Time, d"))+
  theme_bw()
```


```{r}
delta_bc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_cells), group = interaction(Location, Treatment))+
  geom_col(color = "black", fill = "white")+
  geom_errorbar(aes(ymin = ave_delta_cells - sd_delta_cells, ymax = ave_delta_cells + sd_delta_cells), width = 0.1)+
  facet_grid(~factor(Location, levels = levels), scales = "free")+
  labs(x = "", y = expression("Delta Bacterial Carbon, micromol C L"^-1))+
  theme_bw()
```


```{r}
lag <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_lag), group = interaction(Location, Treatment))+
  geom_col(color = "black", fill = "white")+
  geom_errorbar(aes(ymin = ave_lag - sd_lag, ymax = ave_lag + sd_lag), width = 0.1)+
  facet_grid(~factor(Location, levels = levels), scales = "free")+
  labs(x = "", y = "Lag Phase, days")+
  theme_bw()
```

```{r}
#install.packages("patchwork")
library("patchwork")
```


```{r fig.height=6, fig.width=8}
lag + delta_bc + doubling + mew + plot_annotation(tag_levels = "a")
```

