---
title: "Week 4 Notes"
author: "Lexie Christopoulos"
date: "11/6/2020"
output: github_document
---

#Intro 

This document shows how **individual bottle** TOC/DOC data from ACIDD remineralization experiments were processed, QD'd, and analyzed. 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```


#Import Data 

```{r}
excel_sheets("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/ACIDD_Exp_TOC.xlsx")

metadata <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/ACIDD_Exp_TOC.xlsx", sheet = "Metadata")

data <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/ACIDD_Exp_TOC.xlsx",sheet = "Data")

```

```{r}
glimpse(metadata)
```

```{r}
glimpse(data)
```

TOC are samples taken directly from the bottle incubations. 
PTOC stands for parallel TOC, so the incubation vials themselves. 
There are vial samples that go beyond the duration of the sampling for the bottles. 


#Prepare Data 

```{r}
joined <- left_join(metadata, data)

toc <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime)) %>% 
  group_by(Experiment, Bottle) %>%
  mutate(interv = int_length(interval(first(Datetime), Datetime)), hours = interv/3600, days = hours/24) %>%
  ungroup() %>% 
  rename(sd_TOC = TOC_sd, sd_PTOC = PTOC_sd) %>% 
  select(Experiment:Datetime, hours, days, everything(), -c(contains("Sample"), interv)) 

glimpse(toc)



```


# Plot the Curves 

## Pivot Data

```{r}
pivot_toc <- toc %>% 
  select(Experiment, Location, Bottle, Treatment, days, TOC, PTOC) %>% 
  pivot_longer(TOC:PTOC, names_to = "sample", values_to = "value")

pivot_toc_sd <- toc %>% 
  select(Experiment, Location, Bottle, Treatment, days, sd_TOC, sd_PTOC) %>% 
  pivot_longer(sd_TOC:sd_PTOC, names_to = "sample", names_prefix = "sd_", values_to = "sd")

pivoted <- left_join(pivot_toc, pivot_toc_sd) %>% 
  mutate(sample = ifelse(sample == "TOC", "Bottle", "Vial"))

glimpse(pivoted)
```

##Plot 

```{r}
custom.colors <- c("Control" = "#377EB8", "Ash Leachate" = "#4DAF4A", "Santa Barbara" = "#E41A1C", "San Diego" = "#FF7F00", "Vial" = "#377EB8", "Bottle" = "#E41A1C")
levels <- c("Control", "Ash Leachate", "San Diego", "Santa Barbara", "Bottle", "Vial")


pivoted %>% 
  filter(Location == "Santa Barbara") %>% 
  drop_na(value) %>% 
  mutate(Treatment = factor(Treatment, levels = levels),
         sample = factor(sample, levels = levels)) %>% 
  ggplot(aes(x = days, y = value, group = interaction(Treatment, Bottle)))+
  geom_errorbar(aes(ymin = value - sd, ymax = value + sd, color = sample), width = 0.4)+
  geom_point(aes(fill = sample), size = 3, shape = 21, alpha = 0.7)+
  scale_color_manual(values = custom.colors)+
  scale_fill_manual(values = custom.colors)+
  labs(x = "Days", y = expression("Total Organic Carbon, micromol C L"^-1), color = "", fill = "")+
  theme_bw()+
  facet_grid(rows = "Treatment", scales = "free")+
  guides(color = FALSE)

```

The analytical detection limit for our TOC analysis is ~1.5 micromol C L^-1. 


# TOC v PTOC (bottle v vial samples)

```{r}
#install.packages("lmodel2")
library(lmodel2)

reg.data <- toc %>% 
  drop_na(TOC) %>% 
  filter(Timepoint > 0)

reg <- lmodel2(PTOC ~ TOC, data = reg.data, nperm = 99)
```


```{r}
reg
```

```{r}
intercept <- reg$regression.results[3,2]
slope <- reg$regression.results[3,3]

two_int <- reg$confidence.intervals[3,2]
two_slope <- reg$confidence.intervals[3,4]
nine_int <- reg$confidence.intervals[3,3]
nine_slope <- reg$confidence.intervals[3,5]
```

```{r}
reg.data %>% 
  ggplot(aes(x = TOC, y = PTOC))+
  geom_errorbar(aes(ymin = PTOC - sd_PTOC, ymax = PTOC + sd_PTOC), width = 0.05)+
  geom_point(shape = 21, size = 4, alpha = 0.7)+
  geom_abline( intercept = intercept, slope = slope, color = "black", linetype = 2, size = 1)+
  geom_abline(intercept = two_int, slope = two_slope, color = "black", linetype = 3, size = 1)+
  geom_abline(intercept = nine_int, slope = nine_slope, color = "black", linetype = 3, size = 1)+
  labs(x = expression("Bottle TOC, micromol C L"^-1), y = expression("Vial TOC, micromol C L"^-1))+
  theme_bw()+
  annotate(geom = "text", label = expression(atop("y = 0.72x + 19.8", paste("r"^2, "=0.97, ", italic("p "), "=0.01"))), x = 70, y = 74, size = 4)
```


```{r}
bc <- read_rds("C:/Users/aechr/Desktop/github/144l_students/Output_Data/week3/ACIDD_Exp_Processed_BactAbund.rds")

merge <- left_join(toc, bc) %>% 
  select(Experiment:days, TOC:sd_PTOC, cells:diff_ln_cells, bc, ave_bc, sd_bc, everything(), -c(contains("Sample")) ) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  fill(exp_start:sd_lag) %>% 
  ungroup()

glimpse(merge)
```


#Interpolate 

```{r}
#install.packages("zoo")
library(zoo)
```


We only want to interpolate within an experiment and not across experiments, so we'll perform the interpolation by using the split-apply-combine strategy. 

-split the dataframe into a list of its elements (in this case, experiment and bottles)
-apply the interpolate function to each of the list elements 
-combine the results into a new dataframe 


##Split 

```{r}
to_interpolate <- merge %>% 
  select(Experiment, Bottle, Timepoint, days, PTOC, bc) %>% 
  group_by(Experiment, Bottle) #first we'll define the grouping of our dataframe. 

list <- to_interpolate %>% 
  group_split()  #then we convert the dataframe into a list, broken up by the groups (list elements)

keys <- to_interpolate %>% 
  group_keys() %>% 
  mutate(key = paste(Experiment, Bottle))

names(list) <- keys$key
```

##Write the Function 

```{r}
interp.func <- function(x) {
y <- zoo(x, order.by = x$days) #orders our observations 
interp_toc <- round(as.numeric(na.approx(y$PTOC, na.rm = F)), 1) #interpolates toc 
interp_bc <- round(as.numeric(na.approx(y$bc, na.rm = F)), 1)
z <- cbind(y, interp_toc, interp_bc) #combines the columns 
as_tibble(z) #convets to dataframe 
}

```


##Apply and Combine 

```{r}
interpolated <- lapply(list, interp.func) %>% 
  plyr::ldply(., as.data.frame) %>% 
  select(-c(.id, PTOC, bc, days)) %>% 
  mutate_at(vars(Timepoint:interp_bc), as.numeric) %>% 
  left_join(merge, .)

glimpse(interpolated)
```

#Estimate DOC, Bioavalibility, BGEs

```{r}
doc <- interpolated %>% 
  mutate(doc = interp_toc - interp_bc) %>% 
  group_by(Experiment, Bottle) %>% 
  mutate(bioav_doc = (first(doc) - last(na.omit(doc))) / first(doc), 
         doc_exp_end = ifelse(Timepoint == exp_end, doc, NA),
         delta_doc = first(doc) - doc_exp_end,
         tdelta_doc = first(doc) - last(na.omit(doc)),
         bge = ifelse(delta_doc > 1.5, delta_bc/delta_doc, NA)) %>% 
  fill(doc_exp_end:bge, .direction = "downup") %>% 
  ungroup()

glimpse(doc)
```

#Treatment Averages 

```{r}
averages <- doc %>% 
  group_by(Experiment, Treatment, Timepoint) %>% 
  mutate(ave_toc = mean(PTOC),
         sd_toc = sd(PTOC)) %>% 
  ungroup() %>%
  group_by(Experiment, Treatment) %>% 
  mutate(ave_bioav_doc = mean(bioav_doc),
         sd_bioav_doc = sd(bioav_doc),
         ave_delta_doc = mean(delta_doc),
         sd_delta_doc = sd(delta_doc),
         ave_tdelta_doc = mean(tdelta_doc),
         sd_tdelta_doc = sd(tdelta_doc),
         ave_bge = mean(bge),
         sd_bge = sd(bge)) %>% 
  ungroup()


```


# Plot treatment curves (TOC)

```{r}
averages %>% 
  filter(Location == "Santa Barbara") %>% 
  mutate(Treatment = factor(Treatment, levels = levels)) %>% 
  drop_na(ave_toc) %>% 
  ggplot(aes(x = days, y = ave_toc, group = interaction(Experiment, Treatment)))+
  geom_errorbar(aes(ymin = ave_toc - sd_toc, ymax = ave_toc + sd_toc, color = Treatment), width = 0.4)+
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.7)+
  geom_line(aes(color = Treatment), size = 1, alpha = 0.7)+
  scale_color_manual(values = custom.colors)+
  scale_fill_manual(values = custom.colors)+
  labs(x = "Days", y = expression("Total Organic Carbon, micromol C L"^-1), color = "", fill = "")+
  theme_bw()+
  facet_grid(rows = "Treatment", scales = "free")+
  guides(color = FALSE)
```


#Plot treatment averages (bars)

```{r}
bar.data <- averages %>% 
  select(Location, Treatment, ave_bioav_doc:sd_bge) %>% 
  distinct() %>% 
  filter(Location == "Santa Barbara")
```

```{r}
bioav <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_bioav_doc, group = interaction(Location, Treatment)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_bioav_doc - sd_bioav_doc, ymax = ave_bioav_doc + sd_bioav_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Bioavailable DOC Fraction")), color = "")+
  theme_bw()
  
```

```{r}
deldoc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_doc, group = interaction(Location, Treatment)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_delta_doc - sd_delta_doc, ymax = ave_delta_doc + sd_delta_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Delta DOC (to stationary), micromol C L"^-1)), color = "")+
  theme_bw()
```


```{r}
total_deldoc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_tdelta_doc, group = interaction(Location, Treatment)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_tdelta_doc - sd_tdelta_doc, ymax = ave_tdelta_doc + sd_tdelta_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Delta TOC (total), micromol C L"^-1)), color = "")+
  theme_bw()
```

```{r}
bge <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_bge, group = interaction(Location, Treatment)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_bge - sd_bge, ymax = ave_bge + sd_bge), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Bacterial Growth Efficiency")), color = "")+
  theme_bw()
```

```{r}
library(patchwork)
```

```{r fig.height=20, fig.width=10}
deldoc / total_deldoc / bioav / bge + plot_annotation(title = "Santa Barbara Experiments", tag_levels = "a")
```

