---
title: "TOC/DOC Analysis Week 4"
author: "Lexie Christopoulos"
date: "11/6/2020"
output: github_document
---

#Intro 

This document shows how **individual bottle** TOC/DOC data from the 2018  remineralization experiments were processed, QD'd, and analyzed.

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
#install.packages("lmodel2")
library(lmodel2)
#install.packages("zoo")
library(zoo)
#install.packages("patchwork")
library(patchwork)
library(praise)
```


#Import Data 

```{r}
excel_sheets("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx")

metadata <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx", sheet = "Metadata")

data <- read_excel("C:/Users/aechr/Desktop/github/144l_students/Input_Data/week4/144L_2018_Exp_TOC.xlsx",sheet = "Data")
```

```{r}
glimpse(metadata)
glimpse(data)

unique(metadata$Treatment)

```

#Prepare Data 

```{r}
joined <- left_join(metadata, data)

toc <- joined %>% 
  mutate(Datetime = ymd_hm(Datetime)) %>% 
  group_by(Experiment, Bottle) %>%
  mutate(interv = int_length(interval(first(Datetime), Datetime)), hours = interv/3600, days = hours/24) %>%
  ungroup() %>% 
  rename(sd_TOC = TOC_sd, sd_PTOC = PTOC_sd) %>% 
  select(Experiment, Timepoint, Target_DOC_Amendment, Datetime, Bottle, Treatment, DNA_SampleID, TOC:sd_PTOC, hours, days) 

glimpse(toc)
```

#Plot the Curves 

##Pivot Data 

```{r}
pivot_toc <- toc %>% 
  select(Experiment, Bottle, Treatment, days, TOC, PTOC) %>% 
  pivot_longer(TOC:PTOC, names_to = "sample", values_to = "value")

pivot_toc_sd <- toc %>% 
  select(Experiment, Bottle, Treatment, days, DNA_SampleID, sd_TOC, sd_PTOC) %>% 
  pivot_longer(sd_TOC:sd_PTOC, names_to = "sample", names_prefix = "sd_", values_to = "sd")

pivoted <- left_join(pivot_toc, pivot_toc_sd) %>% 
  mutate(sample = ifelse(sample == "TOC", "Bottle", "Vial"))

glimpse(pivoted)
```

##Plot 

```{r}
levels <- c("Control", "Ash Leachate", "Mud Leachate", "Glucose_Nitrate_Phosphate", "Bottle", "Vial")

pivoted %>% 
  drop_na(value) %>% 
  mutate(Treatment = factor(Treatment, levels = levels),
         sample = factor(sample, levels = levels)) %>% 
  ggplot(aes(x = days, y = value, group = interaction(Treatment, Bottle)))+
  geom_errorbar(aes(ymin = value - sd, ymax = value + sd, color = sample), width = 0.4)+
  geom_point(aes(fill = sample), size = 3, shape = 21, alpha = 0.7)+
  labs(x = "Days", y = expression("Total Organic Carbon, micromol C L"^-1), color = "", fill = "")+
  theme_bw()+
  facet_grid(rows = "Treatment", scales = "free")+
  guides(color = FALSE)
```

The analytical detection limit for our TOC analysis is ~1.5 micromol C L^-1. 

#TOC vs PTOC (bottle vs vial samples)

```{r}
reg.data <- toc %>% 
  drop_na(TOC) %>% 
  filter(Timepoint > 0)

reg <- lmodel2(PTOC ~ TOC, data = reg.data, nperm = 99)
```

```{r}
reg
```

```{r}
intercept <- reg$regression.results[3,2]
slope <- reg$regression.results[3,3]

two_int <- reg$confidence.intervals[3,2]
two_slope <- reg$confidence.intervals[3,4]
nine_int <- reg$confidence.intervals[3,3]
nine_slope <- reg$confidence.intervals[3,5]
```


```{r}
reg.data %>% 
  ggplot(aes(x = TOC, y = PTOC))+
  geom_errorbar(aes(ymin = PTOC - sd_PTOC, ymax = PTOC + sd_PTOC), width = 0.05)+
  geom_point(shape = 21, size = 4, alpha = 0.7)+
  geom_abline( intercept = intercept, slope = slope, color = "black", linetype = 2, size = 1)+
  geom_abline(intercept = two_int, slope = two_slope, color = "black", linetype = 3, size = 1)+
  geom_abline(intercept = nine_int, slope = nine_slope, color = "black", linetype = 3, size = 1)+
  labs(x = expression("Bottle TOC, micromol C L"^-1), y = expression("Vial TOC, micromol C L"^-1))+
  theme_bw()+
  annotate(geom = "text", label = expression(atop("y = 1.21x - 2.12", paste("r"^2, "=0.92, ", italic("p "), "=0.01"))), x = 80, y = 90, size = 4)
```


```{r}
bc <- read_rds("C:/Users/aechr/Desktop/github/144l_students/Output_Data/week3/Bacterial_Abundance_Processed_Data.rds")

merge <- left_join(toc, bc) %>% 
  select(Experiment:days,DNA_SampleID, TOC:sd_PTOC, cells:diff_ln_cells, bc, ave_bc, sd_bc, everything()) %>% 
  group_by(Experiment, Treatment, Bottle) %>% 
  fill(exp_start:sd_lag) %>% 
  ungroup()

glimpse(merge)
```

#Interpolate 


We only want to interpolate within an experiment and not across experiments, so we'll perform the interpolation by using the split-apply-combine strategy. 

-split the dataframe into a list of its elements (in this case, experiment and bottles)
-apply the interpolate function to each of the list elements 
-combine the results into a new dataframe

##Split

```{r}
to_interpolate <- merge %>% 
  select(Experiment, Bottle, Timepoint, days, PTOC, bc) %>% 
  group_by(Experiment, Bottle) #first we'll define the grouping of our dataframe. 

list <- to_interpolate %>% 
  group_split()  #then we convert the dataframe into a list, broken up by the groups (list elements)

keys <- to_interpolate %>% 
  group_keys() %>% 
  mutate(key = paste(Experiment, Bottle))

names(list) <- keys$key
```

##Write the Function 
```{r}
interp.func <- function(x) {
y <- zoo(x, order.by = x$days) #orders our observations 
interp_toc <- round(as.numeric(na.approx(y$PTOC, na.rm = F)), 1) #interpolates toc 
interp_bc <- round(as.numeric(na.approx(y$bc, na.rm = F)), 1)
z <- cbind(y, interp_toc, interp_bc) #combines the columns 
as_tibble(z) #convets to dataframe 
}
```

##Apply and Combine 

```{r}
interpolated <- lapply(list, interp.func) %>% 
  plyr::ldply(., as.data.frame) %>% 
  select(-c(.id, PTOC, bc, days)) %>% 
  mutate_at(vars(Timepoint:interp_bc), as.numeric) %>% 
  left_join(merge, .)

glimpse(interpolated)
```

#Estimate DOC, Bioavailibilty and BGEs

```{r}
doc <- interpolated %>% 
  mutate(doc = interp_toc - interp_bc) %>% 
  group_by(Experiment, Bottle) %>% 
  mutate(bioav_doc = (first(doc) - last(na.omit(doc))) / first(doc), 
         doc_exp_end = ifelse(Timepoint == exp_end, doc, NA),
         delta_doc = first(doc) - doc_exp_end,
         tdelta_doc = first(doc) - last(na.omit(doc)),
         bge = ifelse(delta_doc > 1.5, delta_bc/delta_doc, NA)) %>% 
  fill(doc_exp_end:bge, .direction = "downup") %>% 
  ungroup()

glimpse(doc)
```

#Treatment Averages 

```{r}
averages <- doc %>% 
  group_by(Experiment, Treatment, Timepoint) %>% 
  mutate(ave_toc = mean(PTOC),
         sd_toc = sd(PTOC)) %>% 
  ungroup() %>%
  group_by(Experiment, Treatment) %>% 
  mutate(ave_bioav_doc = mean(bioav_doc),
         sd_bioav_doc = sd(bioav_doc),
         ave_delta_doc = mean(delta_doc),
         sd_delta_doc = sd(delta_doc),
         ave_tdelta_doc = mean(tdelta_doc),
         sd_tdelta_doc = sd(tdelta_doc),
         ave_bge = mean(bge),
         sd_bge = sd(bge)) %>% 
  ungroup()

#glimpse(averages)
```

#Plot Treatment Curves (TOC)

```{r}
averages %>% 
  mutate(Treatment = factor(Treatment, levels = levels)) %>% 
  drop_na(ave_toc) %>% 
  ggplot(aes(x = days, y = ave_toc, group = interaction(Treatment, Bottle)))+
  geom_errorbar(aes(ymin = ave_toc - sd_toc, ymax = ave_toc + sd_toc, color = Treatment), width = 0.4)+
  geom_point(aes(fill = Treatment), size = 3, shape = 21, alpha = 0.7)+
  geom_line(aes(color = Treatment), size = 1, alpha = 0.7)+
  labs(x = "Days", y = expression("Total Organic Carbon, micromol C L"^-1), color = "", fill = "")+
  theme_bw()+
  facet_grid(rows = "Treatment", scales = "free")+
  guides(color = FALSE)
```



From these graphs, it looks like all the treatments experianced a decline in TOC. Ash Leachate and the GNP treatment experianced the greatest losses and their curves look almost identical (~ 8 micromoles each). The Mud Leachate treatment only experianced a decline of about 2 micromoles. It started with about the same level of TOC as the Ash Leachate and GNP treatments. The control treatment started with far less TOC than the other treatments, and experianced a decline of about 4 micromoles. 


#Plot Treatment Averages (bars)

```{r}
bar.data <- averages %>% 
  select(Treatment, Bottle, ave_bioav_doc:sd_bge) %>% 
  distinct()
```

###Bioavaliablity 
```{r}
bioav <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_bioav_doc, group = interaction(Treatment, Bottle)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_bioav_doc - sd_bioav_doc, ymax = ave_bioav_doc + sd_bioav_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Bioavailable DOC Fraction")), color = "")+
  theme_bw()
```

###Delta DOC 

```{r}
deldoc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_delta_doc, group = interaction(Treatment, Bottle)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_delta_doc - sd_delta_doc, ymax = ave_delta_doc + sd_delta_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Delta DOC (to stationary), micromol C L"^-1)), color = "")+
  theme_bw()
```

###Total Delta DOC 

```{r}
total_deldoc <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_tdelta_doc, group = interaction(Treatment, Bottle)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_tdelta_doc - sd_tdelta_doc, ymax = ave_tdelta_doc + sd_tdelta_doc), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Delta TOC (total), micromol C L"^-1)), color = "")+
  theme_bw()
```

###Bacterial Growth Efficiency 

```{r}
bge <- bar.data %>% 
  ggplot(aes(x = factor(Treatment, levels = levels), y = ave_bge, group = interaction(Treatment, Bottle)))+
  geom_col(color = "black", fill = "white", alpha = 1)+
  geom_errorbar(aes(ymin = ave_bge - sd_bge, ymax = ave_bge + sd_bge), width = 0.1, size = 0.5)+
  labs(x = "", y = expression(paste("Bacterial Growth Efficiency")), color = "")+
  theme_bw()
```


```{r fig.height=20, fig.width=10}
deldoc / total_deldoc / bioav / bge + plot_annotation(title = "2018 EEMB 144L Experiments", tag_levels = "a")
```


As we can see in graph a, the Delta DOC is highest in the GNP treatment, followed by the Ash Leachate treatment, the control treatment, and the Mud Leachate treatment. In graph b, which represents total Delta DOC, the GNP and Ashe Leachate treatments are equal, at just above 15 micromoles. The control treatment is higher than the Mud Leachate treatment, with around 7 micromoles compared to around 3 micromoles. In graph c, the GNP treatment displays just a little more bioavailable DOC than the Ash Leachate treatment, but they are essentially equal. Once again, the control treatment displays a higher bioavailable DOC than the mud leachate treatment. Graph d shows the bacterial growth efficiency. The control had the highest levels of BGE, followed by the Ash Leachate treatment and the GNP treatment. There appears to be no data for BGE for the Mud Leachate treatment. 


#Save Data 

```{r}
saveRDS(averages, "C:/Users/aechr/Desktop/github/144l_students/Output_Data/week4/TOC_DOC_Analysis_Processed_Data.rds")
```

